### 2022-05-02 src/text/2-package-subsections/analysis.tex
Reformat text files
Fix pre-commit hooks

The Analysis class has the responsability of orchestrating of the workflow defined in the configuration (AnalysisConfig) and triggering the execution of the AnalysisStep classes that defines the common use cases identified. These steps include the following: 
 observations selection from the DataStore,  data reduction, excess map computation, model fitting, flux-points estimation, and light-curves production. The structure of the Analysis class allows to define custom AnalysisStep,  so ones could design and configure arbitrarily complex wokflows.

>>>>

The Analysis class has the responsibility of orchestrating of the workflow
defined in the configuration (AnalysisConfig) and triggering the execution of
the AnalysisStep classes that defines the common use cases identified. These
steps include the following: observations selection from the DataStore,  data
reduction, excess map computation, model fitting, flux-points estimation, and
light-curves production. The structure of the Analysis class allows to define
custom AnalysisStep,  so ones could design and configure arbitrarily complex
wokflows.

### 2022-05-02 src/text/2-package-subsections/irf.tex
Reformat text files
Fix pre-commit hooks

\item Energy dispersion: Edisp, in \iact, is generally given in terms of the so called migration parameter ($\mu$), which is defined as the ratio between the reconstructed energy
and the true energy. This ratio should be as close as one and its dispersion can assume the shape of a gaussian (or even more complex distributions). Migration parameter
is given at each offset angle and reconstructed energy. The main sub-classes are {\it EnergyDispersion2D}, designed to interpret Edisp, {\it EDispKernelMap}, which builds an
Edisp kernel map, i.e. a 4-dimensional WCS map where at each sky-position is associated an Edisp kernel. The latter is a representation of the Edisp as a function of the true energy only thanks to the sub-class {\it EDispKernel}.

\item Background: the BKG can be represented in \gammapy as either 1) a 2D map ({\it Background2D}) of count rate normalised per steradians and energy at different energies
and offset-angles or 2) as rate per steradians and energy, as a function of reconstructed energy and detector coordinates ({\it Background3D}). In the former, the background is expected to follow a radially symmetric shape, while in the latter, it can be more complex.
\end{itemize}

>>>>

	\item Energy dispersion: Edisp, in \iact, is generally given in terms of the so
	      called migration parameter ($\mu$), which is defined as the ratio between the
	      reconstructed energy and the true energy. This ratio should be as close as one
	      and its dispersion can assume the shape of a gaussian (or even more complex
	      distributions). Migration parameter is given at each offset angle and
	      reconstructed energy. The main sub-classes are {\it EnergyDispersion2D},
	      designed to interpret Edisp, {\it EDispKernelMap}, which builds an Edisp kernel
	      map, i.e., a 4-dimensional WCS map where at each sky-position is associated an
	      Edisp kernel. The latter is a representation of the Edisp as a function of the
	      true energy only thanks to the sub-class {\it EDispKernel}.

	\item Background: the BKG can be represented in \gammapy as either 1) a 2D map ({\it
			      Background2D}) of count rate normalised per steradians and energy at different
	      energies and offset-angles or 2) as rate per steradians and energy, as a
	      function of reconstructed energy and detector coordinates ({\it Background3D}).
	      In the former, the background is expected to follow a radially symmetric shape,
	      while in the latter, it can be more complex.

\end{itemize}

### 2022-05-02 src/text/2-package-subsections/stats.tex
Reformat text files
Fix pre-commit hooks

The gammapy.stats subpackage contains the fit statistics and associated statistical estimators
that are commonly used in gamma-ray astronomy.
In general, gamma-ray observations count Poisson-distributed events at various sky positions,
and contain both signal and background events. Estimation of the number of signal events is done
through likelihood maximization. In Gammapy, the fit statistics are Poisson log-likelihood functions
normalized like chi-squares, i.e. they follow the expression $2 \times log L$, where $L$ is the
likelihood function used.
The statistic function used when the expected number of background events is known
is the \emph{Cash} statistic ~\citep{Cash}. It is used by datasets using background
templates such as the MapDataset. When the number of background events is unknown and
an off measurement where only background events are expected is used, the statistic function
is WStat. It is a profile log-likelihood statistic where background counts are
marginalized parameters. It is used by datasets containing off counts measurements
such as the SpectrumDatasetOnOff, used for classical spectral analysis.

To perform simple statistical estimations on counts measurements, CountsStatistic classes
encapsulate the aforementioned statistic functions to measure excess counts and estimate the
associated statistical significance, errors and upper limits. They perform maximum likelihood
ratio tests to estimate significance (the square root of the statistic difference) and compute
likelihood profiles to measure errors and upper limits. The following example allows to compute
the Li \& Ma significance~\citep{LiMa} of a set of measurements:

>>>>

The gammapy.stats subpackage contains the fit statistics and associated
statistical estimators that are commonly used in gamma-ray astronomy. In
general, gamma-ray observations count Poisson-distributed events at various sky
positions, and contain both signal and background events. Estimation of the
number of signal events is done through likelihood maximization. In Gammapy,
the fit statistics are Poisson log-likelihood functions normalized like
chi-squares, i.e., they follow the expression $2 \times log L$, where $L$ is
the likelihood function used. The statistic function used when the expected
number of background events is known is the \emph{Cash} statistic
~\citep{Cash}. It is used by datasets using background templates such as the
MapDataset. When the number of background events is unknown and an off
measurement where only background events are expected is used, the statistic
function is WStat. It is a profile log-likelihood statistic where background
counts are marginalized parameters. It is used by datasets containing off
counts measurements such as the SpectrumDatasetOnOff, used for classical
spectral analysis.

To perform simple statistical estimations on counts measurements,
CountsStatistic classes encapsulate the aforementioned statistic functions to
measure excess counts and estimate the associated statistical significance,
errors and upper limits. They perform maximum likelihood ratio tests to
estimate significance (the square root of the statistic difference) and compute
likelihood profiles to measure errors and upper limits. The code example
\ref{codeexample:stats} shows how to compute the Li \& Ma
significance~\citep{LiMa} of a set of measurements.

\begin{figure}
	\import{code-examples/generated/}{gp_stats}
	\caption{}
	\label{codeexample:stats} \end{figure}

### 2022-05-02 src/text/3-applications-subsections/multi-instrument-analysis.tex
Fix pre-commit hooks

In this multi-instrument analysis example we showcase the capabilities of
\gammapy to perform a simultaneous likelihood fit incorporating data from
different instruments and at different levels of reduction. We estimate the
spectrum of the Crab Nebula combining data from \fermi, \magic and \hawc. Maps
of \fermi data are prepared selecting a region of $X^{\circ}$ around the
position of the Crab Nebula applying the same selection criteria of the 3FHL
catalog (7 years of data with energy from $10\,{\rm GeV}$ to $2\,{\rm TeV}$,
\citealt{3FHL}).  The \magic data are two%% shall we explain exactly which type of Map is created?
observations of $20\,{\rm min}$ each, chosen from the dataset used to estimate
the performance of the upgraded stereo system \citep{magic_performance} and
already included in \cite{joint_crab}. The observations were taken at small
zenith angles ($<30^{\circ}$) in wobble mode \citep{fomin_1994}, with the
source sitting at an offset of $0.4^{\circ}$ from the FoV center. Their energy
range spans $80\,{\rm GeV} -- 20\,{\rm TeV}$. They are reduced to ON/OFF
dataset before being fitted. \hawc flux points data are estimated in
\cite{hawc_crab_2019} with $2.5\,{\rm years}$ of data and span an energy range
$300\,{\rm GeV} -- 300\,{\rm TeV}$, and directly read with \gammapy.

>>>>

In this multi-instrument analysis example we showcase the capabilities of
\gammapy to perform a simultaneous likelihood fit incorporating data from
different instruments and at different levels of reduction. We estimate the
spectrum of the Crab Nebula combining data from \fermi, \magic and \hawc. Maps
of \fermi data are prepared selecting a region of $X^{\circ}$ around the
position of the Crab Nebula applying the same selection criteria of the 3FHL
catalog (7 years of data with energy from $10\,{\rm GeV}$ to $2\,{\rm TeV}$,
~\citealt{3FHL}).  The \magic data are two%% shall we explain exactly which type of Map is created?
observations of $20\,{\rm min}$ each, chosen from the dataset used to estimate
the performance of the upgraded stereo system~\citep{magic_performance} and
already included in~\cite{joint_crab}. The observations were taken at small
zenith angles ($<30^{\circ}$) in wobble mode~\citep{fomin_1994}, with the
source sitting at an offset of $0.4^{\circ}$ from the FoV center. Their energy
range spans $80\,{\rm GeV} -- 20\,{\rm TeV}$. They are reduced to ON/OFF
dataset before being fitted. \hawc flux points data are estimated
in~\cite{hawc_crab_2019} with $2.5\,{\rm years}$ of data and span an energy
range $300\,{\rm GeV} -- 300\,{\rm TeV}$, and directly read with \gammapy.

### 2022-05-02 src/text/3-applications-subsections/population-studies.tex
Fix pre-commit hooks

\subsection{Surveys, catalogs, and population studies}
\label{ssec:population-studies}

Sky surveys have a large potential for new source detections, and new phenomena
discovery. They also offer less selection bias to perform source population
studies over a large set of coherently detected and modelled objects. Early
versions of \textit{gammapy} were developed in parallel of the preparation of
the HESS Galactic plane survey catalog \citep[HGPS][]{2018A&A...612A...1H} and
the associated PWN and SNR populations studies \citep{2018A&A...612A...2H,
	2018A&A...612A...3H}.

The increase in sensitivity  and resolution provided by new generation of
instruments scale up the number of sources detectable and the complexity of the
models needed to represent them accurately. As an example If we compare the
results of the HGPS to the expectations from the CTA Galactic Plane survey
simulations we jump from 78 sources detected by HESS to about 500 detectable by
CTA  \citep{2021arXiv210903729R}.

Studies performed on simulations not only offer a first glimpse on what could
be the sky seen by CTA (according to our current knowledge on source
populations), but also give us the opportunity to test the software on complex
use cases\footnote{Note that the CTA-GPS simulations were performed with the
	\textit{ctools} package \citep{2016A&A...593A...1K} and analysed with both
	\textit{ctools} and \textit{gammapy} packages in order to cross-validate
	them.}. So we can  improve performances, optimize our analyses strategies, and
identify the needs in term of parallelisation to process the large datasets
provided by the surveys.

In short the production of catalogs from gamma-ray surveys can be divided in
four main steps : data-reduction; object detection; model fitting and model
selection; associations and classification. The IACTs data-reduction step is
done in the same way than described in the previous sections but scale-up to
few thousand of observations. The object detection step consists a minima in
finding local maxima in the significance, or TS maps,  given by the
ExcessMapEstimator, or TSMapEstimator, respectively.  Further refinements can
include for example  filtering				and detection on these
maps with techniques from scikit-image package	\citep{scikit-image}, and
outlier detection from scikit-learn package \citep{scikit-learn}. Tests on
simulations shown that it reduces the spurious detections at this stage and
then speed up the next step as less objects will have to be fitted
simultaneously. During the modelling step each object is alternatively fitted
with different models in order to determine their optimal parameters, and the
best-candidate model. The subpackage gammapy.modeling.models offers a large
variety of choice, and the possibility to add custom models.  Several spatial
models (point-source, disk, gaussian...), and spectral models (power-law,
log-parabola...) may be tested for each object, so the complexity of the
problem increases rapidly in regions crowded with multiple extended sources.
Finally an object is discarded if its best-fit model is not significantly
preferred over the null hypothesis (no source) comparing the difference in
log-Likelihood between these two hypotheses. For the association and
classification step, that is tightly connected to the population studies, we
can a minima compare the fitted models to the set of gammapy-ray catalogs
available in gammapy.catalog. However further multi-wavelength cross-matches
are usually required to characterize the sources.

>>>>

\subsection{Surveys, catalogs, and population studies}
\label{ssec:surveys-catalogs-and-population-studies}

Sky surveys have a large potential for new source detections, and new phenomena
discovery. They also offer less selection bias to perform source population
studies over a large set of coherently detected and modelled objects. Early
versions of \textit{gammapy} were developed in parallel of the preparation of
the HESS Galactic plane survey catalog~\citep[HGPS][]{2018A&A...612A...1H} and
the associated PWN and SNR populations studies~\citep{2018A&A...612A...2H,
	2018A&A...612A...3H}.

The increase in sensitivity  and resolution provided by new generation of
instruments scale up the number of sources detectable and the complexity of the
models needed to represent them accurately. As an example If we compare the
results of the HGPS to the expectations from the CTA Galactic Plane survey
simulations we jump from 78 sources detected by HESS to about 500 detectable by
CTA~\citep{2021arXiv210903729R}.

Studies performed on simulations not only offer a first glimpse on what could
be the sky seen by CTA (according to our current knowledge on source
populations), but also give us the opportunity to test the software on complex
use cases\footnote{Note that the CTA-GPS simulations were performed with the
	\textit{ctools} package~\citep{2016A&A...593A...1K} and analysed with both
	\textit{ctools} and \textit{gammapy} packages in order to cross-validate
	them.}. So we can  improve performances, optimize our analyses strategies, and
identify the needs in term of parallelisation to process the large datasets
provided by the surveys.

In short the production of catalogs from gamma-ray surveys can be divided in
four main steps : data-reduction; object detection; model fitting and model
selection; associations and classification. The IACTs data-reduction step is
done in the same way than described in the previous sections but scale-up to
few thousand of observations. The object detection step consists a minima in
finding local maxima in the significance, or TS maps,  given by the
ExcessMapEstimator, or TSMapEstimator, respectively.  Further refinements can
include for example  filtering and detection on these maps with techniques from
scikit-image package~\citep{scikit-image}, and outlier detection from
scikit-learn package~\citep{scikit-learn}. Tests on simulations shown that it
reduces the spurious detections at this stage and then speed up the next step
as less objects will have to be fitted simultaneously. During the modelling
step each object is alternatively fitted with different models in order to
determine their optimal parameters, and the best-candidate model. The
subpackage gammapy.modeling.models offers a large variety of choice, and the
possibility to add custom models.  Several spatial models (point-source, disk,
gaussian...), and spectral models (power-law, log-parabola...) may be tested
for each object, so the complexity of the problem increases rapidly in regions
crowded with multiple extended sources. Finally an object is discarded if its
best-fit model is not significantly preferred over the null hypothesis (no
source) comparing the difference in log-Likelihood between these two
hypotheses. For the association and classification step, that is tightly
connected to the population studies, we can a minima compare the fitted models
to the set of gammapy-ray catalogs available in gammapy.catalog. However
further multi-wavelength cross-matches are usually required to characterize the
sources.

### 2022-05-02 src/text/3-applications-subsections/temporal-analysis.tex
Reformat text files
Fix pre-commit hooks

A common use case in most astrophysical scenarios is to study the temporal variability of the source. The most basic way to do this is to construct a lightcurve, ie, the flux of the source in each given time bin. In gammapy, this is done by using the LightCurveEstimator which fits the normalisation of the source in each energy band per observation, keeping other parameters constant. For custom time binning, an observation needs to be split into finer ones using the \verb"Observation.select_time method". Figure \ref{fig:hess_lightcurve_pks} shows the lightcurve of the blazar PKS~2155-304 in different energy bands as observed by the H.E.S.S. telescope during an exceptional flare on the night of July 29 - 30, 2006 \cite{2009A&A...502..749A}. The data is available publicly as a part of the HESS-DL3-DR1 \cite{HESS-DL3-DR1}, and shipped with \verb"GAMMAPY_DATA". Each observation is first split into 10 min smaller observations, and spectra extracted for each of these within a 0.11 deg radius around the source. A PowerLawSpectralModel is fit to all the datasets, leading to a reconstructed index of $3.54 \pm 0.02$. With this assumed spectral model the \verb"LightCurveEstimator" run directly for two energy bands, 0.5 - 1.5 TeV, and 1.5 - 20 TeV,  respectively.

>>>>

A common use case in most astrophysical scenarios is to study the temporal
variability of the source. The most basic way to do this is to construct a
lightcurve, ie, the flux of the source in each given time bin. In gammapy, this
is done by using the LightCurveEstimator which fits the normalisation of the
source in each energy band per observation, keeping other parameters constant.
For custom time binning, an observation needs to be split into finer ones using
the \verb"Observation.select_time method". Figure \ref{fig:hess_lightcurve_pks}
shows the lightcurve of the blazar PKS~2155-304 in different energy bands as
observed by the H.E.S.S. telescope during an exceptional flare on the night of
July 29 - 30, 2006~\cite{2009A&A...502..749A}. The data is available publicly
as a part of the HESS-DL3-DR1~\cite{HESS-DL3-DR1}, and shipped with
\verb"GAMMAPY_DATA". Each observation is first split into 10 min smaller
observations, and spectra extracted for each of these within a 0.11 deg radius
around the source. A PowerLawSpectralModel is fit to all the datasets, leading
to a reconstructed index of $3.54 \pm 0.02$. With this assumed spectral model
the \verb"LightCurveEstimator" run directly for two energy bands, 0.5 - 1.5
TeV, and 1.5 - 20 TeV,	respectively.

### 2022-05-02 src/text/4-project.tex
Fix pre-commit hooks

\section{Gammapy project}
\label{sec:gammapy-project}

@@

Several automatized tasks are set as GitHub
actions\footnote{\url{https://github.com/features/actions}}, blocking the
processes and alerting developers when fails occur. This is the case of the
content integration workflow, which monitors the execution of the test coverage
suite\footnote{\url{https://pytest.org}} using datasets from the
\textit{gammapy-data} repository. Tests scan not only the codebase but also the
code snippets present in docstrings of the scripts and in the RST documentation
files, as well as in the tutorials provided in the form of Jupyter notebooks.

Other automatized tasks, executing in the \textit{gammapy-benchmarks}
repository, are responsible for numerical validation tests and benchmarks
monitoring. Also tasks related with the release process are partially
automatized, and every contribution to the codebase repository triggers the
documentation building and publishing workflow within the \textit{gammapy-docs}
repository (see Sec. ~\ref{ssec:software-distribution} and Sec.
~\ref{ssec:documentation}).

This small ecosystem of interconnected up-to-date repositories, automatized
tasks and alerts, is just a part of a bigger set of GitHub repositories, where
most of them are related with the project but not necessary for the development
of the software (i.e. project webpage, complementary high-energy astrophysics
object catalogs, coding sprints and weekly developer calls minutes,
contributions to conferences, other digital assets, etc.) Finally, third-party
services for code quality metrics are also set and may be found as status
shields in the codebase repository.

@@

\subsection{Roadmap}
\todo{Axel, Regis}

>>>>

# Custom rule to prepare multi-instrument datasets
rule prepare_multi_instrument: input:
"src/data/multi-instrument/input/fermi/Fermi-LAT-3FHL_data_Fermi-LAT.fits",
"src/data/multi-instrument/input/fermi/Fermi-LAT-3FHL_iem.fits",
"src/data/multi-instrument/input/fermi/Fermi-LAT-3FHL_datasets.yaml",
"src/data/multi-instrument/input/fermi/Fermi-LAT-3FHL_models.yaml",
"src/data/multi-instrument/input/magic/hdu-index.fits.gz",
"src/data/multi-instrument/input/magic/obs-index.fits.gz",
"src/data/multi-instrument/input/magic/20131004_05029747_DL3_CrabNebula-W0.40+035.fits",
"src/data/multi-instrument/input/magic/20131004_05029748_DL3_CrabNebula-W0.40+215.fits",
"src/data/multi-instrument/input/hawc/HAWC19_flux_points.fits", output:
"src/data/multi-instrument/datasets/flux_points/crab_magic_flux_points.fits",
"src/data/multi-instrument/datasets/flux_points/crab_fermi_flux_points.fits",
"src/data/multi-instrument/results/crab_multi_instrument_fit.yaml", conda:
"environment.yml" shell: "cd src/data/multi-instrument && python make.py"
\section{Gammapy project} \label{sec:gammapy-project}

@@

Several automated tasks are set as GitHub
actions\footnote{\url{https://github.com/features/actions}}, blocking the
processes and alerting developers when fails occur. This is the case of the
content integration workflow, which monitors the execution of the test coverage
suite\footnote{\url{https://pytest.org}} using datasets from the
\textit{gammapy-data} repository. Tests scan not only the codebase but also the
code snippets present in docstrings of the scripts and in the RST documentation
files, as well as in the tutorials provided in the form of Jupyter notebooks.

Other automated tasks, executing in the \textit{gammapy-benchmarks} repository,
are responsible for numerical validation tests and benchmarks monitoring. Also
tasks related with the release process are partially automated, and every
contribution to the codebase repository triggers the documentation building and
publishing workflow within the \textit{gammapy-docs} repository (see Sec.
~\ref{ssec:software-distribution} and Sec. ~\ref{ssec:documentation}).

This small ecosystem of interconnected up-to-date repositories, automated tasks
and alerts, is just a part of a bigger set of GitHub repositories, where most
of them are related with the project but not necessary for the development of
the software (i.e., project webpage, complementary high-energy astrophysics
object catalogs, coding sprints and weekly developer calls minutes,
contributions to conferences, other digital assets, etc.) Finally, third-party
services for code quality metrics are also set and may be found as status
shields in the codebase repository.

@@

\subsection{Roadmap}
\label{ssec:roadmap}
\todo{Axel, Regis}
